---
title: "French presidential election poll tracker"
author: "DorotaW"
date: "18 April 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
```


## French elections - who is leading in the first round?

```{r}
library(plotly)
library(xlsx)
library(tidyr)
library(dplyr)

polls <- read.xlsx("Election Data.xlsx", sheetIndex = 1)
events <- read.xlsx("Election Data.xlsx", sheetIndex = 2)
polls$SAMPLE <- as.integer(polls$SAMPLE)
events$y <- 0.01

polls_avg <- polls[2:9] %>%
  group_by (Datetime) %>% 
  summarise_each (funs(mean))

polls_pivot <- gather(polls_avg, "Candidate", "Perc", 3:7)
polls_pivot$Candidate <- ordered(as.factor(polls_pivot$Candidate))
levels(polls_pivot$Candidate) <- c("François Fillon", "Benoît Hamon", "Marine Le Pen", "Jean-Luc Mélenchon", "Emmanuel Macron")

ymax=max(polls_pivot$Perc)
line_list <- list()

for(i in 1:nrow(events)){ 
     line_list[[i]] <- 
          list(type      = "line",
               line      = list(color = 'black', width = 0.5, dash = 'dot'),
               x0        = events$Datetime[i],
               x1        = events$Datetime[i],
               xref      = "x",
               y0        = 0, 
               y1        = ymax,
               yref      = "y")
}

plot_ly() %>%
  add_trace(data=polls_pivot, x = ~Datetime, y = ~Perc, 
            color = ~Candidate,  colors = "Set1",
            text = ~paste0(Candidate, ": ", sprintf("%1.0f%%", 100*Perc)), hoverinfo = 'text',
            type = "scatter", mode = "lines", 
            inherit = F, showlegend = T) %>%
  add_trace(data=events, x = ~Datetime, y = ~y, 
            text = ~paste0(Datetime, ": ", Event), hoverinfo = 'text',
            marker =  list(color = "black", symbol = "star", size = 16),
            type = "scatter", mode = "markers",
            inherit = F, showlegend = F) %>% 
  layout(hovermode='closest',
         shapes = line_list,
         legend = list(orientation = 'h'),
         xaxis = list(showgrid=F , zeroline=F, showline=T , ticks="", title="Date", 
                      range = c("2017-01-01","2017-04-23")),
         yaxis = list(showgrid=F , zeroline=F, showline=T , ticks="inside", title="Voting Intention (%)",
                      tickformat=".0%", range = c(0, 0.30)))

```

*Source: [https://ig.ft.com](https://ig.ft.com/sites/france-election/polls/)*

## Can we trust these results?

The tracker was created using national opinions polls.

Given fine polling margins and sampling methods, those kind of trackers are anything but conclusive. Political polling might not be the best way of predicting exact electoral outcomes, but they provide a trend of public sentiment towards the candidates.
