---
title: "What impacts miles per gallon - analysis of mtcars data set"
author: "Dorota"
date: "21 June 2015"
output: word_document
---

## Executive Summary

Based on below analysis,  we can coclude that manual transmissions are better than automatic in terms of miles per gallon.
A change from automatic to manual transmission increased the MPG by 7.245. However, transmission type only explained 36% (R2) of the variation in mpg. Other variables like weight, number of cylinders and horsepower are better indicator of miles per gallon.


## Data Processing

After uploading the data, some of the vriables were coded as factor
```{r,results='hide'}
mtcars<-mtcars
mtcars$cyl=factor(mtcars$cyl)
mtcars$vs=factor(mtcars$vs)
mtcars$gear=factor(mtcars$gear)
mtcars$carb=factor(mtcars$carb)
mtcars$am=factor(mtcars$am,labels=c('Automatic','Manual'))
```

## Exploratory Data Analysis**

Below boxplot shows that there might be a difference between automatic and manual  transmission

```{r}
library(ggplot2)
ggplot(mtcars, aes(x=am, y = mpg, fill=am)) +
  geom_boxplot() + labs(title = "MIles per gallon")
```

```{r,echo=FALSE}
library(plyr)

test=t.test(mpg ~am, data = mtcars)
t=t.test(mpg ~am, data = mtcars)$p.value
```

Additionally, running t test shows that this difference is significant (t-test results in p-value `r t`)


```{r,echo=FALSE}
library(plyr)
### mean of transmission type
Transmission=ddply(mtcars,"am",summarize, avg=mean(mpg))
test=t.test(mpg ~am, data = mtcars)
t=t.test(mpg ~am, data = mtcars)$p.value
```

## Building the model

We wil start building the model, buy using only tranmission variable, later on we will use stepwise approach to find the best model (based on AIC criterion)

```{r}
library(MASS)
fit<- lm(mpg ~ am, data=mtcars)
fit2<-lm(mpg ~ ., data=mtcars)
summary(fit)
step = stepAIC(fit2, scope=list(lower=~am),direction="both",trace=FALSE)
summary(step)
```

Adding number of cylinders, weight and horsepower to the model improves how much variance can be explained by the model (from 36% to 87% ). Also coefficient changed from 7.245 to 1.8


## Diagnostics

We can also check residuals

``` {r}
par(mfrow=c(2,2))
plot(fit)
```

## Conclusion

There is a difference in mpg in relation to transmission type, however, transmission type does not appear to be a very good explanatory variable on its own.  
